---
title: "survXgboost package"
author: "Iyar Lin"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, cache = F, eval = T)
set.seed(1)
options(scipen = 999)
```

The xgboost package survival model returns predictions on the hazard ratio scale 
(i.e., as $HR = exp(marginal_prediction) in the proportional hazard function 
$h(t) = h0(t) * HR$. This quantity is equivalent to the type = "risk" in coxph. 
This package provides a thin wrapper that enables using the xgboost package to 
perform full survival curve estimation.

See also [discussion](https://datascience.stackexchange.com/questions/65266/how-do-i-predict-survival-curves-using-xgboost) in stackoverflow overflow

Below is a short usage demo.

First, prepare the data:

```{r}
pacman::p_load_gh("IyarLin/survXgboost")
pacman::p_load("survival")
pacman::p_load("xgboost")
data("lung")
lung <- lung[complete.cases(lung), ] # doesn't handle missing values at the moment
lung$status <- lung$status - 1 # format status variable correctly such that 1 is event/death and 0 is censored/alive
label <- ifelse(lung$status == 1, lung$time, -lung$time)

val_ind <- sample.int(nrow(lung), 0.1 * nrow(lung))
x_train <- as.matrix(lung[-val_ind, !names(lung) %in% c("time", "status")])
x_label <- label[-val_ind]
x_val <- xgb.DMatrix(as.matrix(lung[val_ind, !names(lung) %in% c("time", "status")]),
                     label = label[val_ind])
```

Below we train an xgboost survival model using the function from the survXgboost
package rather than the xgboost package:

```{r}
# train surv_xgboost
surv_xgboost_model <- xgb.train.surv(
  params = list(
    objective = "survival:cox",
    eval_metric = "cox-nloglik",
    eta = 0.05 # larger eta leads to algorithm not converging, resulting in NaN predictions
  ), data = x_train, label = x_label,
  watchlist = list(val2 = x_val),
  nrounds = 1000, early_stopping_rounds = 30
)
```

Next we can predict full survival curves:

```{r}
# predict survival curves
times <- seq(10, 1000, 50)
survival_curves <- predict(object = surv_xgboost_model, newdata = x_train, type = "surv", times = times)
matplot(times, t(survival_curves[1:5, ]), type = "l")
```

We can also predict the risk scores like in the original xgboost package:

```{r}
# predict risk score
risk_scores <- predict(object = surv_xgboost_model, newdata = x_train, type = "risk")
hist(risk_scores)
```


